
```{r setup, include=FALSE, cache=FALSE}
library(knitr)
# opts_knit$set(out.format = "latex")
knit_theme$set(knit_theme$get("greyscale0"))

options(replace.assign = FALSE, width = 50)

opts_chunk$set(fig.path = "figure/graphics-", 
               cache.path = "cache/graphics-", 
               fig.align = "center", 
               dev = "pdf", fig.width = 5, fig.height = 5, 
               fig.show = "hold", cache = FALSE, par = TRUE)
knit_hooks$set(crop = hook_pdfcrop)

knit_hooks$set(par = function(before, options, envir) {
  if (before && options$fig.show != "none") {
    #par(mar=c(3,3,2,1),cex.lab=.95,cex.axis=.9, mgp=c(2,.7,0),tcl=-.01, las=1)
  }}, crop = hook_pdfcrop)
```

The aim of this practical is to review our knowledge of functions, for loops and if statements. 

# Basic functions

Consider the following simple function
```{r echo=TRUE}
v = 5
Fun1 = function() {
  v = 0
  return(v)
}
Fun1()
``` 

1. Why does the final line return 0 and not 5.
    ```{r }
    ## Fun1 uses the local variable v
    ```
1. Delete line 3 in the above piece of code. Now change `Fun1()` to allow
`v` to be passed as an argument, i.e. we can write `Fun1(5)`. Call
this function to make sure it works.

    ```{r}
    Fun1 = function(v) {
      return(v)
    }
    Fun1(10)
    ```


## Default arguments:

```{r echo=TRUE}
Fun2 = function(x = 10) {
  return(x)
}

Fun3 = function(x) {
  return(x)
}
``` 

1. Why does 
    ```{r echo=TRUE}
    Fun2()
    ```
    work, but this raises an error
    ```{r eval=FALSE, echo=TRUE}
    Fun3()
    ```

    ```{r}
    ## Fun3 expects an argument x, but 
    ## we haven't given one and there is
    ## no default.
    ```

1. Change `Fun2` so that it returns `x*x`.

    ```{r }
    Fun2 = function(x = 10) {
      return(x * x)
    }
    ``` 

## `if` statements.

```{r echo=TRUE}
Fun4 = function(x) {
  if (x == 5) {
    y = 0
  } else {
    y = 1
  }
  return(y)
}
``` 

Change `Fun4` so that it:

  * returns 1 if $x$ is positive;
  * returns -1 if $x$ is negative;
  * returns 0 if $x$ is zero.

    ```{r }
    Fun4 = function(x) {
      rtn_value = 0
      if (x > 0) {
        rtn_value = 1
      } else if (x < 0) {
        rtn_value = -1
      }
      return(rtn_value)
    }
    ```   

Change `Fun4()` so it errors if $x$ is positive

    ```{r }
    Fun4 = function(x) {
      rtn_value = 0
      if (x > 0) {
        stop("This function requires x <= 0")
      } else if (x < 0) {
        rtn_value = -1
      }
      return(rtn_value)
    }
    ```   

## `for` loops.

```{r results="hide", echo=TRUE}
total = 0
for (i in 1:5) {
  total = total + i
}
total
``` 

The `for` loop above calculates 
\[
\sum_{i=1}^5 i = 1 + 2 + 3 + 4 + 5
\]

1. What is the final value of `total` in the above piece of code?
    ```{r }
    total
    ```
1. Change the above loop to calculate the following summations:
    \[
\mbox{(i)} \sum_{i=1}^{20} (i+1)  
\]

    ```{r }
    total = 0
    for (i in 1:20) {
      total = total + (i + 1)
    }
    total
    ```
    
\[
\mbox{(ii)} \sum_{j=-10}^{15} j 
\]

    ```{r }
    total = 0
    for (j in-10:15) {
      total = total + j
    }
    total
    ```
1. Rewrite the two loops using the `sum()` function. For example, the for loop in the first
example can be written as `sum(1:5)`

    ```{r}
    sum(2:21)
    sum(-10:15)
    ```

## More functions, for loops and signalling conditions:

```{r echo=TRUE}
a = 2
total = 0
for (blob in a:5) {
  total = total + blob
}
```

1. In the code above, delete line 1. Now put the above code in a function called `Fun5`,
where `a` is passed as an argument, i.e. we can call
`Fun5(1)`

    ```{r }
    Fun5 = function(a) {
      total = 0
      for (blob in a:5) {
        total = total + blob
      }
      return(total)
    }
    Fun5(1)
    ```

1. Alter the code so that the `for` loop goes from `a` to
`b`, rather than `a` to $5$. Allow `b` to be passed as an
argument, i.e. we can call `Fun5(1,5)`.
    ```{r }
    Fun5 = function(a, b) {
      total = 0
      for (blob in a:b) {
        total = total + blob
      }
      return(total)
    }
    Fun5(1, 5)
    ```

1. Change `Fun5` so that it has default arguments of
`a = 1` and `b = 10`.

    ```{r }
    Fun5 = function(a = 1, b = 10) {
      total = 0
      for (blob in a:b) {
        total = total + blob
      }
      return(total)
    }
    Fun5(5)
    ``` 
    
1. Change `Fun5` so that it messages the user the total after each iteration and stops the function if the total has surpassed 50.

    ```{r , message = FALSE}
    Fun5 = function(a = 1, b = 10) {
      total = 0
      for (blob in a:b) {
        total = total + blob
        message("Current total is ", total)
        if (total > 50) {
          stop("Total has surpassed limit of 50")
        }
      }
      return(total)
    }
    Fun5(5)
    ```

## Solutions

The solutions can be viewed via

```{r eval=FALSE, echo = TRUE}
library("jrAdvPackage")
vignette("solutions1", package = "jrAdvPackage")
```

